---
title: "IIT Admission Trend (2016-2022"
author: "Kamleshwar"
date: "2023-06-12"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project
I analyse admission trend, not only of India's but worlds best insitute IIT from 2016 to 2022

## Reading data

```{r cars}
library(dplyr)
library(ggplot2)
library(plotrix)
setwd("C:/Users/Admin/Downloads/archive/")
csv<- read.csv("student_data.csv",header = TRUE, sep = ",")
csv_tbl<- as_tibble(csv)
head(csv_tbl)
#to see all column heading
colnames(csv)
```

## Creating new column
we create new column that we use for analysis
```{r}
#adding new column in dataset
discount_percentage<- (csv$Discount.on.Fees/csv$Fees)*100
discount_percentage<- round(discount_percentage,2)
csv$discount_percentage<-(discount_percentage)
#again checking all column heading
colnames(csv)
```

## Admision per year
Now we check admission taken per year in IIT
```{r}
admission_per_year<-(count(group_by(csv,Year.of.Admission)))
print(admission_per_year)
```

now we plot the graph

## Including Plots
```{r, echo=FALSE}
ggplot(admission_per_year, aes(x = Year.of.Admission, y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "IIT Admission Trend (2016-2022)",
       x = "Year of Admission", y = "Number of Admissions") +
  geom_text(aes(label = n), hjust = 0.5, vjust = 20, color = "white") +
  theme_minimal()
```

as we can see admission taken per year is almost around at 28,000

## Poular Department
Now we check which department have most student
```{r}
field_with_most_students<-(count(group_by(csv,Field.of.Study)))
print(field_with_most_students)
```

lets plot graph

```{r,echo=FALSE}
ggplot(field_with_most_students, aes(x = Field.of.Study, y = n)) +
  geom_bar(stat = "identity", fill = "grey") +
  labs(title = "IIT Admission Trend (2016-2022)",
       x = "Branch", y = "Number of Admissions") +
  geom_text(aes(label = n), hjust = 0.5, vjust = 20, color = "white") +
  theme_minimal()
```

now we check pie chart but firt we create a data frame that calculate percentage

```{r}
department_percentage= (field_with_most_students$n/sum(field_with_most_students$n))*100
department_percentage=round(department_percentage,2)
department_percentage<-data.frame(Department = field_with_most_students$Field.of.Study, Percentage = department_percentage)
department_percentage
```

now let see pie chart

```{r,echo=FALSE}
pie3D(department_percentage$Percentage,
      labels = department_percentage$Percentage,
      col = rainbow(length(department_percentage$Percentage)))
legend("bottomright", department_percentage$Department,
       cex = 0.5, fill = rainbow(length(department_percentage$Percentage)))

```

## Comparing Fess and Discount
```{r,echo=FALSE}
# Create scatter plot
pairs(~Fees + Discount.on.Fees, data = csv,
      col.axis = "darkgreen")

```

now we check maximum and minimun fees and discount respectivly but first we have to create a dat frame

```{r}
data <- data.frame(
  category = c("Fees", "Discount"),
  value = c(max(csv$Fees), max(csv$Discount.on.Fees)),
  ymin = c(min(csv$Fees), min(csv$Discount.on.Fees)),
  ymax = c(max(csv$Fees), max(csv$Discount.on.Fees))
)
```

```{r,echo=FALSE}
ggplot(data, aes(x = category, y = value)) +
  geom_bar(stat = "identity", fill = "gray") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 1.5, color = "red") +
  labs(title = "Maximum and Minimum Fees and Discounts",
       x = "Category", y = "Value") +
  theme_minimal()

```

